Effect of quality-adjusted production function inputs

### 2015-05-15 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Now using cleaner code to make all models.

<<Formulas>>=
# Set up the formulas for each type of analysis.
# These formulas contain energy. 
# We can use keep_summands to drop the energy term (3rd summand on the rhs)
# later if needed.
formulas <- 
  list( 
    PT = c(unadjusted = build_formula("iGDP", list("iKstkS.L", "iL", "iXpMP", "iYear")),
             adjusted = build_formula("iGDP", list("iKservS.L", "ihLest", "iUMP", "iYear"))),
    UK = c(unadjusted = build_formula("iGDP", list("iKservO.WwithRD", "iL", "iXp", "iYear")),
             adjusted = build_formula("iGDP", list("iKservO.WwithRD", "ihLest", "iU", "iYear")))
  )
@

@@

<<Historical Data, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Historical data.", warning=FALSE>>=
histDataForGraph <- leaf_apply(formulas, class = "formula", f=function(form, name){
  meta <- strsplit(name, split="[.]") %>% unlist
  Country <- meta[1]
  flavor <- meta[2]
  measure.vars <- c(summands(form, right=FALSE), summands(form, 1:3))
  trimmed_historical_data <- historicalData[[Country]][c("Year", measure.vars)]
  mtrimmed_historical_data <- melt(trimmed_historical_data, measure.vars = measure.vars, 
                                   value.name = "value", variable.name = "variable")
  mtrimmed_historical_data$Country <- Country
  if (flavor == "unadjusted"){
    mtrimmed_historical_data$flavor <- "Unadjusted"
  } else if (flavor == "adjusted") {
    mtrimmed_historical_data$flavor <- "Quality-adjusted"
  } else {
    stop(paste("Unknown flavor", flavor, "in histDataForGraph."))
  }
  return(mtrimmed_historical_data)
}) %>% do.call(rbind, .)

@@

<<Pre-econometric graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Pre-econometric data.", warning=FALSE>>=

# Create some pre-econometric ratios

# # Melt the data
# mdata_preeconometric <- melt(historicalData, 
#                              measure.vars = c("kPerGDP.unadj", "lPerGDP.unadj", "ePerGDP.unadj",
#                                               "kPerGDP.adj", "lPerGDP.adj", "ePerGDP.adj"),
#                              variable.name = "variable",
#                              value.name = "value")
# # Add a column to say whether the variable is Unadjusted or Quality-adjusted based on variable name
# mdata_preeconometric$adjusted <- lapply(mdata_preeconometric$variable, function(x) {
#   suffix <- unlist(strsplit(as.character(x), split="[.]")) [2]
#   if (suffix == "unadj"){
#     return("Unadjusted")
#   } else if (suffix == "adj"){
#     return("Quality-adjusted")
#   } else {
#     stop(paste("Unknown suffix:", suffix))
#   }
# })

pre_econometric_data <- leaf_apply(formulas, class = "formula", f=function(form, name){
  meta <- strsplit(name, split="[.]") %>% unlist
  Country <- meta[1]
  flavor <- meta[2]
  nums <- summands(form, n=1:3)
  denom <- summands(form, n=1, right=FALSE)
  ratios <- lapply(nums, function(num) {paste(num, denom, sep="/")}) %>% unlist
  coltemplate <- build_formula(NULL, c("Year", ratios))
  trimmed_historical_data <- historicalData[[Country]][all.vars(coltemplate)]
  ratio_data <- evalFormula(coltemplate, data=trimmed_historical_data)$right
  mratio_data <- melt(ratio_data, measure.vars = ratios, value.name = "ratio", variable.name = "variable")
  mratio_data$Country <- Country
  mratio_data$flavor <- flavor
  return(mratio_data)
}) %>% do.call(rbind, .)

@@

<<Models>>=
fileName <- "ESEE2015Models.Rdata"
if (!rerunModels){
  # load from the file on disk
  load(fileName)
} else {
  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.
  nest <- NULL
  nest$kl <- c(1,2)
  nest$kle <- c(1,2,3)
  nest$lek <- c(2,3,1)
  nest$ekl <- c(3,1,2)

#   models <- list()
#   for ()

#   models <- NULL
#   # Portugal, CD
#   models$PT$unadj$noE$CD <- cdModel(formula = f$PT$unadj$noE, 
#                                     data = dataPT, constrained = TRUE)
#   models$PT$unadj$withE$CD <- cdModel(formula = f$PT$unadj$withE, 
#                                       data = dataPT, constrained = TRUE)
#   models$PT$adj$noE$CD <- cdModel(formula = f$PT$adj$noE, 
#                                   data = dataPT, constrained = TRUE)
#   models$PT$adj$withE$CD <- cdModel(formula = f$PT$adj$withE, 
#                                     data = dataPT, constrained = TRUE)
#   # UK, CD
#   models$UK$unadj$noE$CD <- cdModel(formula = f$UK$unadj$noE, 
#                                     data = dataUK, constrained = TRUE)
#   models$UK$unadj$withE$CD <- cdModel(formula = f$UK$unadj$withE, 
#                                       data = dataUK, constrained = TRUE)
#   models$UK$adj$noE$CD <- cdModel(formula = f$UK$adj$noE, 
#                                   data = dataUK, constrained = TRUE)
#   models$UK$adj$withE$CD <- cdModel(formula = f$UK$adj$withE, 
#                                     data = dataUK, constrained = TRUE)
#   # PT, CESkl
#   models$PT$unadj$noE$CESkl <- cesModel(formula = f$PT$unadj$noE, 
#                                         nest = nest$kl, data = dataPT, constrained = TRUE)
#   models$PT$adj$noE$CESkl <- cesModel(formula = f$PT$adj$noE, 
#                                       nest = nest$kl, data = dataPT, constrained = TRUE)
#   # PT, CESkle
#   models$PT$unadj$withE$CESkle <- cesModel(formula = f$PT$unadj$withE, 
#                                            nest = nest$kle, data = dataPT, constrained = TRUE)
#   models$PT$adj$withE$CESkle <- cesModel(formula = f$PT$adj$withE, 
#                                          nest = nest$kle, data = dataPT, constrained = TRUE)
#   # PT, CESlek
#   models$PT$unadj$withE$CESlek <- cesModel(formula = f$PT$unadj$withE, 
#                                            nest = nest$lek, data = dataPT, constrained = TRUE)
#   models$PT$adj$withE$CESlek <- cesModel(formula = f$PT$adj$withE, 
#                                          nest = nest$lek, data = dataPT, constrained = TRUE)
#   # PT, CESekl
#   models$PT$unadj$withE$CESekl <- cesModel(formula = f$PT$unadj$withE, 
#                                            nest = nest$ekl, data = dataPT, constrained = TRUE)
#   models$PT$adj$withE$CESekl <- cesModel(formula = f$PT$adj$withE, 
#                                          nest = nest$ekl, data = dataPT, constrained = TRUE)
#   # UK, CESkl
#   models$UK$unadj$noE$CESkl <- cesModel(formula = f$UK$unadj$noE, 
#                                         nest = nest$kl, data = dataUK, constrained = TRUE)
#   models$UK$adj$noE$CESkl <- cesModel(formula = f$UK$adj$noE, 
#                                       nest = nest$kl, data = dataUK, constrained = TRUE)
#   # UK, CESkle
#   models$UK$unadj$withE$CESkle <- cesModel(formula = f$UK$unadj$withE, 
#                                            nest = nest$kle, data = dataUK, constrained = TRUE)
#   models$UK$adj$withE$CESkle <- cesModel(formula = f$UK$adj$withE, 
#                                          nest = nest$kle, data = dataUK, constrained = TRUE)
#   # UK, CESlek
#   models$UK$unadj$withE$CESlek <- cesModel(formula = f$UK$unadj$withE, 
#                                            nest = nest$lek, data = dataUK, constrained = TRUE)
#   models$UK$adj$withE$CESlek <- cesModel(formula = f$UK$adj$withE, 
#                                          nest = nest$lek, data = dataUK, constrained = TRUE)
#   # UK, CESekl
#   models$UK$unadj$withE$CESekl <- cesModel(formula = f$UK$unadj$withE, 
#                                            nest = nest$ekl, data = dataUK, constrained = TRUE)
#   models$UK$adj$withE$CESekl <- cesModel(formula = f$UK$adj$withE, 
#                                          nest = nest$ekl, data = dataUK, constrained = TRUE)
  # Save models to disk.
  save(models, file=fileName)
}
@

>>>>

<<Formulas>>=
# Set up the formulas for each type of analysis.
# These formulas contain energy. 
# We can use keep_summands to drop the energy term (3rd summand on the rhs)
# later if needed.
formulas <- 
  list( 
    PT = c(unadjusted = build_formula("iGDP", list("iKstkS.L", "iL", "iXpMP", "iYear")),
             adjusted = build_formula("iGDP", list("iKservS.L", "ihLest", "iUMP", "iYear"))),
    UK = c(unadjusted = build_formula("iGDP", list("iKservO.WwithRD", "iL", "iXp", "iYear")),
             adjusted = build_formula("iGDP", list("iKservO.WwithRD", "ihLest", "iU", "iYear")))
  )
countries <- names(formulas)
@

@@

<<Historical Data, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Historical data.", warning=FALSE>>=
histDataForGraph <- leaf_apply(formulas, class = "formula", f=function(form, name){
  meta <- strsplit(name, split="[.]") %>% unlist
  Country <- meta[1]
  flavor <- meta[2]
  measure.vars <- c(summands(form, right=FALSE), summands(form, 1:3))
  trimmed_historical_data <- historicalData[[Country]][c("Year", measure.vars)]
  mtrimmed_historical_data <- melt(trimmed_historical_data, measure.vars = measure.vars, 
                                   value.name = "value", variable.name = "variable")
  mtrimmed_historical_data$Country <- Country
  mtrimmed_historical_data$flavor <- if (flavor == "adjusted") "Quality-adjusted" else "Unadjusted"
  return(mtrimmed_historical_data)
}) %>% do.call(rbind, .)

@@

<<Pre-econometric graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Pre-econometric data.", warning=FALSE>>=
pre_econometric_data <- leaf_apply(formulas, class = "formula", f=function(form, name){
  meta <- strsplit(name, split="[.]") %>% unlist
  Country <- meta[1]
  flavor <- meta[2]
  nums <- summands(form, n=1:3)
  denom <- summands(form, n=1, right=FALSE)
  ratios <- lapply(nums, function(num) {paste(num, denom, sep="/")}) %>% unlist
  coltemplate <- build_formula(NULL, c("Year", ratios))
  trimmed_historical_data <- historicalData[[Country]][all.vars(coltemplate)]
  ratio_data <- evalFormula(coltemplate, data=trimmed_historical_data)$right
  mratio_data <- melt(ratio_data, measure.vars = ratios, value.name = "ratio", variable.name = "variable")
  mratio_data$Country <- Country
  mratio_data$flavor <- if (flavor == "adjusted") "Quality-adjusted" else "Unadjusted"
  return(mratio_data)
}) %>% do.call(rbind, .)

@@

<<Models>>=
fileName <- "ESEE2015Models.Rdata"
if (!rerunModels){
  # load from the file on disk
  load(fileName)
} else {
  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.
  flavors <- c("unadjusted", "adjusted")
  nests <- NULL
  nests$kle <- c(1,2,3); nests$lek <- c(2,3,1); nests$ekl <- c(3,1,2)
  energies <- c(FALSE, TRUE)

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, data = historicalData[[country]], nest=c(1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- 
        lapply(nests, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest)
        })
    }      
  }
}
  # Save models to disk.
  save(models, file=fileName)
}
@
